# Figure 4
Validation of the RF8 model

## Validation of RF8 model (RJAML-Cohort 2)

```{r}
# Load the trained model
# rf_tuned <- readRDS("../03.out/figure3/AZA_VEN_RandomForest8.rds")
# 
# # Load the new data for Cohort 2
# newDat <- read_xlsx("../02.data/rjaml_test_all.xlsx")
# newDat <- newDat[, -1]  # Remove the first column (Patient ID)
# 
# # Select features and log-transform
# data_selected <- newDat[, c("Response", selected_features)]
# data_selected[, -1] <- log2(data_selected[, -1] + 1)
# 
# # Predict class labels and probabilities for Cohort 2
# new_predictions <- predict(rf_tuned, newdata = data_selected)
# new_probabilities <- predict(rf_tuned, newdata = data_selected, type = "prob")
# 
# # Add predictions to the dataset
# newDat$Predicted_Response <- new_predictions
# newDat$RF8.prob.CR <- new_probabilities[, "CR"]
# write.csv(newDat, "../03.out/figure4/RF8_Cohort2.csv")
```

```{r}
# Compute and plot ROC for Cohort 2
newDat <- read.csv("../03.out/figure4/RF8_Cohort2.csv", header = T)
roc_obj <- roc(newDat$Response, newDat$RF8.prob.CR)
roc_obj$auc

plot(smooth(roc_obj), legacy.axes = TRUE, lwd = 2,
     main = "Cohort 2 (N = 32)",
     xlab = "False Positive Rate (1 - Specificity)", 
     ylab = "True Positive Rate (Sensitivity)", 
     cex.main = 1.5, cex.lab = 1.2, cex.axis = 1.1)
abline(h = seq(0, 1, by = 0.1), v = seq(0, 1, by = 0.1), col = "gray80", lty = 2)
legend("bottomright", legend = c(paste("AUC =", round(roc_obj$auc, 2))),
       col = "#1f77b4", lwd = 2, bty = "n", cex = 1.2)

# Compute and plot Precision-Recall (PR) curve for Cohort 2
newDat$Response <- factor(newDat$Response, levels = c("NonCR", "CR"))
response_binary <- as.numeric(newDat$Response) - 1
pr_curve <- pr.curve(scores.class0 = newDat$RF8.prob.CR, weights.class0 = response_binary, curve = TRUE)

plot(pr_curve, main = "Cohort 2 (N = 32)", color = "#1f77b4", auc.main = FALSE)
abline(h = seq(0, 1, by = 0.1), v = seq(0, 1, by = 0.1), col = "gray80", lty = 2)
legend("bottomright", legend = c(paste("AUPRC =", round(pr_curve$auc.integral, 2))),
       col = "#1f77b4", lwd = 2, bty = "n", cex = 1.2)

```

## Validation of RF8 model (FPMTB)

```{r}
# newDat <- read_xlsx("../02.data/Validation_Exvivo_FPMTB_84Genes_43R_55S.xlsx")
# newDat <- newDat[, -1]  # Remove the first column (Patient ID)
# # Select features and log-transform
# data_selected <- newDat[, c("Response", selected_features)]
# data_selected[, -1] <- log2(data_selected[, -1] + 1)
# 
# # Predict class labels and probabilities for the FPMTB set
# new_predictions <- predict(rf_tuned, newdata = data_selected)
# new_probabilities <- predict(rf_tuned, newdata = data_selected, type = "prob")
# 
# # Add predictions to the dataset
# newDat$Predicted_Response <- new_predictions
# newDat$RF8.prob.CR <- new_probabilities[, "CR"]
# write.csv(newDat, "../03.out/figure4/RF8_FPMTB.csv")
```

```{r}
newDat <- read.csv("../03.out/figure4/RF8_FPMTB.csv", header = T)
roc_obj <- roc(newDat$Response, newDat$RF8.prob.CR)
roc_obj$auc

plot(smooth(roc_obj), legacy.axes = TRUE, lwd = 2,
     main = "FPMTB set (N = 98)",
     xlab = "False Positive Rate (1 - Specificity)", 
     ylab = "True Positive Rate (Sensitivity)", 
     cex.main = 1.5, cex.lab = 1.2, cex.axis = 1.1)
abline(h = seq(0, 1, by = 0.1), v = seq(0, 1, by = 0.1), col = "gray80", lty = 2)
legend("bottomright", legend = c(paste("AUC =", round(roc_obj$auc, 2))),
       col = "#1f77b4", lwd = 2, bty = "n", cex = 1.2)

# Compute and plot Precision-Recall (PR) curve for the FPMTB set
newDat$Response <- factor(newDat$Response, levels = c("NonCR", "CR"))
response_binary <- as.numeric(newDat$Response) - 1
pr_curve <- pr.curve(scores.class0 = newDat$RF8.prob.CR, weights.class0 = response_binary, curve = TRUE)

plot(pr_curve, main = "FPMTB set (N = 98)", color = "#1f77b4", auc.main = FALSE)
abline(h = seq(0, 1, by = 0.1), v = seq(0, 1, by = 0.1), col = "gray80", lty = 2)
legend("bottomright", legend = c(paste("AUPRC =", round(pr_curve$auc.integral, 2))),
       col = "#1f77b4", lwd = 2, bty = "n", cex = 1.2)
```

## Validation of RF8 model (BeatAML)

```{r}
# newDat <- read_xlsx("../02.data/Validation_Exvivo_BeatAML_84Genes_170R_88S.xlsx")
# newDat <- newDat[, -1]  # Remove the first column (Patient ID)
# # Select features and log-transform
# data_selected <- newDat[, c("Response", selected_features)]
# data_selected[, -1] <- log2(data_selected[, -1] + 1)
# # Predict class labels and probabilities for the BeatAML set
# new_predictions <- predict(rf_tuned, newdata = data_selected)
# new_probabilities <- predict(rf_tuned, newdata = data_selected, type = "prob")
# # Add predictions to the dataset
# newDat$Predicted_Response <- new_predictions
# newDat$RF8.prob.CR <- new_probabilities[, "CR"]
# write.csv(newDat, "../03.out/figure4/RF8_BeatAML.csv")
```

```{r}
newDat <- read.csv("../03.out/figure4/RF8_BeatAML.csv", header = T)
# Compute and plot ROC for the BeatAML set
roc_obj <- roc(newDat$Response, newDat$RF8.prob.CR)
roc_obj$auc

plot(smooth(roc_obj), legacy.axes = TRUE, lwd = 2,
     main = "BeatAML set (N = 258)",
     xlab = "False Positive Rate (1 - Specificity)", 
     ylab = "True Positive Rate (Sensitivity)", 
     cex.main = 1.5, cex.lab = 1.2, cex.axis = 1.1)
abline(h = seq(0, 1, by = 0.1), v = seq(0, 1, by = 0.1), col = "gray80", lty = 2)
legend("bottomright", legend = c(paste("AUC =", round(roc_obj$auc, 2))),
       col = "#1f77b4", lwd = 2, bty = "n", cex = 1.2)


# Compute and plot Precision-Recall (PR) curve for the BeatAML set
newDat$Response <- factor(newDat$Response, levels = c("NonCR", "CR"))
response_binary <- as.numeric(newDat$Response) - 1
pr_curve <- pr.curve(scores.class0 = newDat$RF8.prob.CR, weights.class0 = response_binary, curve = TRUE)

plot(pr_curve, main = "BeatAML set (N = 258)", color = "#1f77b4", auc.main = FALSE)
abline(h = seq(0, 1, by = 0.1), v = seq(0, 1, by = 0.1), col = "gray80", lty = 2)
legend("bottomright", legend = c(paste("AUPRC =", round(pr_curve$auc.integral, 2))),
       col = "#1f77b4", lwd = 2, bty = "n", cex = 1.2)

```

## Violin and Boxplots for Response Probability (RF8) 

```{r}
library(ggplot2)
library(reshape2)
library(ggpubr)
library(patchwork)

# Define a function to generate plots for each dataset
generate_plot <- function(file_path, dataset_label) {
  data <- read.csv(file_path, header = TRUE, row.names = 1)
  data$dataset <- dataset_label
  
  p <- ggplot(data, aes(x = Response, y = RF8.prob.CR, fill = Response)) +
    geom_violin(trim = FALSE) +
    geom_boxplot(width = 0.1, position = position_dodge(0.9), outlier.shape = NA) +
    scale_fill_manual(values = c("#93C89A", "#FFCC98")) +
    facet_wrap(.~ dataset) +
    labs(title = "",
         x = "True Response",
         y = "Response probability (RF8)") +
    theme_classic() +
    theme(legend.position = "none") +
    stat_compare_means(aes(label = ..p.format..), comparisons = list(c("CR", "NonCR")))
  
  return(p)
}

p6 <- generate_plot("../03.out/figure4/RF8_Cohort2.csv", "Cohort 2")
p7 <- generate_plot("../03.out/figure4/RF8_FPMTB.csv", "FPMTB set")
p8 <- generate_plot("../03.out/figure4/RF8_BeatAML.csv", "BeatAML set")

plots <- p6 | p7 | p8 
plots

```

## RF8 score among clusters

```{r}
library(ggplot2)
library(reshape2)
library(ggpubr)
library(patchwork)

dat1 <- read.csv("../03.out/figure4/RF8_FPMTB.csv", header = T)
p1 <- ggplot(dat1, aes(x = factor(Cluster), y = RF8.prob.CR, fill = factor(Cluster))) +
    geom_violin(trim = FALSE, alpha = 0.5) +
    geom_boxplot(width = 0.25, alpha = 0.75, position = position_dodge(0.9), outlier.shape = NA) +
    labs(title = "FPMTB", x = "Cluster", y = "Response probability (RF8)") +
    theme_classic() +
    theme(legend.position = "none") +
    stat_compare_means(
        comparisons = list(c('1', '2'), c('1', '3'), c('2', '3')),
        method = "wilcox.test",
        label = "p.signif"
    ) +
    scale_fill_manual(values = c("#e9b383", "#7c9d97", "#9cb0c3"))
p1

dat2 <- read.csv("../03.out/figure4/RF8_BeatAML.csv", header = T)
p2 <- ggplot(dat2, aes(x = factor(Cluster), y = RF8.prob.CR, fill = factor(Cluster))) +
    geom_violin(trim = FALSE, alpha = 0.5) +
    geom_boxplot(width = 0.25, alpha = 0.75, position = position_dodge(0.9), outlier.shape = NA) +
    labs(title = "BeatAML", x = "Cluster", y = "Response probability (RF8)") +
    theme_classic() +
    theme(legend.position = "none") +
    stat_compare_means(
        comparisons = list(c('1', '2'), c('1', '3'), c('2', '3')),
        method = "wilcox.test",
        label = "p.signif"
    ) +
    scale_fill_manual(values = c("#e9b383", "#7c9d97", "#9cb0c3"))
p2

ggsave(p1, filename = "../03.out/figure4/RF8_FPMTB_clusters.pdf", width = 5, height = 3.6)
ggsave(p2, filename = "../03.out/figure4/RF8_BeatAML_clusters.pdf", width = 5, height = 3.6)

```


## Integration of ex vivo drug-response in BeatAML

```{r eval=TRUE, message = FALSE, warning = FALSE}
# Obtain the clinically approved drugs for AML
library(reshape2)
library(ggridges)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(ggpubr)
library(patchwork)
drug_auc <- read.table("../02.data/01_BeatAML_Inhibitor_AUC_values_v4_dbgap.txt", header = T,sep = "\t", check.names = FALSE)
drug_select <- subset(drug_auc, inhibitor %in% c("Azacytidine", "Venetoclax"))
drug_select <- drug_select[, c("Sample", "inhibitor", "auc")]
dat <- dcast(drug_select, Sample~inhibitor)
colnames(dat) <- c("Sample", "Azacitidine", "Venetoclax" )
id <- read.csv("../02.data/02_RNAseq_Sample_ids.csv", header = T, check.names = FALSE) 
dat <- merge(dat, id, by = "Sample")
row.names(dat) <- dat$Sample
dat <- subset(dat, dat$diseaseStageAtSpecimenCollection %in% c("Initial Diagnosis", "Relapse"))
dat <- subset(dat, dat$rnaSeq == "y")
### select samples having more than 2 drugs tested
dat$NA_number <- rowSums(is.na(dat[, 2:3]))
dat <- subset(dat, NA_number == 0)
dat <- dat[, c("Azacitidine", "Venetoclax")]
head(dat)

```

## BeatAML Waves 1-4 heatmap based on drug profiling

```{r}
library(pheatmap)
core_mat <- dat
core_mat1 <- as.matrix(t(core_mat))
core_mat1 <- log2(core_mat1 + 1)
# core_mat1 <- core_mat1 - apply(core_mat1, 1, mean)
# core_mat1[which(core_mat1 < -2)] <- -2
# core_mat1[which(core_mat1 > 2)] <- 2
bk <- unique(c(seq(-2, 2, length=100)))
out <- pheatmap(core_mat1,
                breaks = bk, 
                cutree_cols = 2,
                color = colorRampPalette(c("#8C510A", "#F5F5F5",  "#01665E"))(100), 
                scale = "row", border_color = "black", angle_col = 90,
                show_rownames = TRUE, show_colnames = FALSE,
                clustering_method = "ward.D",
                #annotation_colors = ann_colors,
                #annotation_col = annotation_col,
                cluster_cols = T, cluster_rows = T
)
# Order of clustered samples
clust_sample <- colnames(core_mat1[,out$tree_col[["order"]]])
clust_sample <- data.frame(clust_sample)
sample_cluster <- data.frame(cutree(out$tree_col, k=3))
colnames(sample_cluster) <- "Cluster"
data <- data.frame(core_mat)
data <- cbind(data, sample_cluster)
data$group <- ifelse(data$Cluster == 1, "Resistant", "Sensitive")
table(data$group)
# write.csv(data, "../03.out/figure4/BeatAML_exVivo_Resistant_Sensitive_Raw.csv")

sample_cluster$Lable <- ifelse(data$Cluster == 1, "Resistant", "Sensitive")
annotation_col = data.frame(
  Lable = factor(sample_cluster$Lable)
  )
rownames(annotation_col) = colnames(core_mat1)
ann_colors = list(
  Lable = c(Sensitive = "#7c9d97", Resistant = "#e9b383")
)

# pdf("../03.out/figure4/01.BeatAML_exVivo_drug_heatmap.pdf", width = 10, height = 2)
out <- pheatmap(core_mat1,
                breaks = bk, 
                cutree_cols = 3,
                color = colorRampPalette(c("#8C510A", "#F5F5F5",  "#01665E"))(100), 
                scale = "row", border_color = "black", angle_col = 90,
                show_rownames = TRUE, show_colnames = FALSE,
                clustering_method = "ward.D",
                annotation_colors = ann_colors,
                annotation_col = annotation_col,
                cluster_cols = T, cluster_rows = T
)
# dev.off()

# Extract scaled values from the heatmap object
# Create matrix and scale manually (same scaling as pheatmap uses)
core_matrix <- as.matrix(core_mat1)
scale_rows <- function(x) {
  m <- apply(x, 1, mean, na.rm = TRUE)
  s <- apply(x, 1, sd, na.rm = TRUE)
  return((x - m) / s)
}
scaled_matrix <- scale_rows(core_matrix)
# Get cluster assignments (k=3)
cluster_assignments <- data.frame(Cluster = cutree(out$tree_col, k = 3))
# Combine scaled values with cluster assignments
result_data <- cbind(t(scaled_matrix), cluster_assignments)
# write.csv(result_data, "../03.out/figure4/BeatAML_exVivo_Resistant_Sensitive_scaled_scores.csv")
```

## Boxplot for drug sensitivitity between resistant and sensitive

```{r}
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(tidyverse)
library(ggrastr)
library(ggbeeswarm)
mat <- read.csv("../03.out/figure4/BeatAML_exVivo_Resistant_Sensitive_Raw.csv", header = T, row.names = 1)
mat <- mat[, c("group", "Azacitidine", "Venetoclax")]
mat <- mat %>% gather(-"group", key = "drug", value = "AUC") 
mat$drug <- factor(mat$drug, levels = c("Azacitidine","Venetoclax"))
p <- ggplot(mat, aes(x = group, y = AUC, color = group)) + 
  geom_boxplot_jitter() + geom_quasirandom(width=0.15, alpha = 0.75) +
  facet_wrap(.~drug, ncol=3) + 
  theme_pubr() + 
  xlab("") + ylab("Drug resistance (AUC)") +
  theme(strip.text.x = element_text(size=13)) +
  stat_compare_means(label = "p.signif", comparisons = list(c('Resistant', 'Sensitive')), method='wilcox.test') + 
  scale_color_manual(values = c("#e9b383", "#7c9d97")) + 
  theme(legend.position='None')
p

## Cluster (1 2 3)
mat <- read.csv("../03.out/figure4/BeatAML_exVivo_Resistant_Sensitive_Raw.csv", header = T, row.names = 1)
mat <- mat[, c("Azacitidine", "Venetoclax", "Cluster")]
mat <- mat %>% gather(-"Cluster", key = "drug", value = "sDSS") 

mat$Cluster <- as.factor(mat$Cluster)

p1 <- ggplot(mat, aes(x = Cluster, y = sDSS, color = Cluster)) + 
  geom_boxplot_jitter() + 
  geom_quasirandom(width=0.15, alpha = 0.5) +
  facet_wrap(.~drug) + 
  theme_pubr() + 
  xlab("") + ylab("Drug resistance (AUC)") +
  theme(strip.text.x = element_text(size=13)) +
  stat_compare_means(comparisons = list(c('1', '2'), c('1', '3'), c('2', '3')), 
                     method='wilcox.test', label =  "p.signif") + 
  scale_color_manual(values = c("#e9b383", "#7c9d97", "#9cb0c3")) +  
  theme(legend.position='None')
p1

###################################
####### Scaled sensitivity scores
###################################

# Clusters (1, 2, 3)
drug_sensitivity_cluster <- read.csv("../03.out/figure4/BeatAML_exVivo_Resistant_Sensitive_scaled_scores.csv", header = TRUE, row.names = 1)
drug_sensitivity_cluster <- drug_sensitivity_cluster[, c("Azacitidine", "Venetoclax", "Cluster")]
drug_sensitivity_cluster <- gather(drug_sensitivity_cluster, -Cluster, key = "drug", value = "sDSS")

drug_sensitivity_cluster$Cluster <- as.factor(drug_sensitivity_cluster$Cluster)

p2 <- ggplot(drug_sensitivity_cluster, aes(x = Cluster, y = sDSS, color = Cluster)) + 
  geom_boxplot_jitter() + geom_quasirandom(width = 0.15, alpha = 0.5) +
  facet_wrap(~drug) + 
  theme_pubr() + 
  xlab("") + ylab("Scaled Drug sensitivity (sDSS)") +
  theme(strip.text.x = element_text(size = 13)) +
  stat_compare_means(comparisons = list(c('1', '2'), c('1', '3'), c('2', '3')), method = 'wilcox.test', label = "p.signif") + 
  scale_color_manual(values = c("#e9b383", "#7c9d97", "#9cb0c3")) +  
  theme(legend.position = 'None')

p2
```

